---
title: "Interative Plot Exercises"
output:
  html_document:
    df_print: paged
    code_folding: hide
---


```{r intro, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(ggplot2)
library(data.table)
library(plotly)
library(tidyquant)
library(leaflet)
library(magrittr)
diamonds <- diamonds[sample(x = nrow(diamonds),size = nrow(diamonds)*0.1),]
```

### 1. PLOT THIS GRAPH
```{r ggplot, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(diamonds,aes(x=color,y=cut)) + geom_jitter(aes(alhpa=clarity, color=factor(color)))
```

### 2. MAKE IT DYNAMIC
```{r plotly, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(diamonds,aes(x=color,y=cut)) + geom_jitter(aes(alhpa=clarity,color=factor(color))) -> gg
ggplotly(gg)
```



### 3. GRAB A TIME SERIE
##### get the informatoin of AAPL stocks by using  tq_get() function from tidyquant package
```{r tidyquant, message=FALSE, warning=FALSE, paged.print=FALSE}
tidyquant::tq_get('AAPL')
```

### 4. TIBBLE TO XTS
##### use the function  tbl_xts() function from tbl2xts package
```{r tbl2xts, message=FALSE, warning=FALSE, paged.print=FALSE}
tbl2xts::tbl_xts(tq_get('AAPL')) -> xts
head(xts)
```


### 5. DYGRAPH
##### create the first dygraph using the funciton dygraph() from dygraphs package
```{r dygraph, message=FALSE, warning=FALSE, paged.print=FALSE}
dygraphs::dygraph(xts)
```

### 6. PLOT THE ADJUSTED PRICE INSTEAD OF THE VOLUME
##### add a range selector, a highlighter and a mean roller 
##### find out how to do it [here](https://rstudio.github.io/dygraphs/)
```{r dygraph_1, message=FALSE, warning=FALSE, paged.print=FALSE}
dygraphs::dygraph(xts[,'adjusted'])
```


### 7. ADD WIDGETS TO THE PREVIOUS DYGRAPH
##### add a range selector, a highlighter and a mean roller 
##### find out how to do it [here](https://rstudio.github.io/dygraphs/)
```{r dygraph_2, message=FALSE, warning=FALSE, paged.print=FALSE}
dygraphs::dygraph(xts[,'adjusted']) %>% 
  dygraphs::dyRangeSelector() %>% 
  dygraphs::dyHighlight() %>% 
  dygraphs::dyRoller()

```


### 8. GEOCODE A TEXT
##### find the longitude and latitude of "IE BUSINESS SCHOOL" by using the function gecode() from the package ggmap
```{r geo, message=FALSE, warning=FALSE, paged.print=FALSE}
ggmap::geocode('IE BUSINESS SCHOOL')
```


### 9. GEOCODE A TEXT
##### find the longitude and latitude of "IE BUSINESS SCHOOL" by using the function gecode() from the package ggmap
```{r geo_1, message=FALSE, warning=FALSE, paged.print=FALSE}
geo <- ggmap::geocode('IE BUSINESS SCHOOL')
leaflet() %>% addTiles() %>% addMarkers(lat = as.numeric(geo$lat),
                                        lng = as.numeric(geo$lon)) 
```

### 10. CHANGE THE BASEMAP
##### find out [here](https://rstudio.github.io/leaflet/basemaps.html) how to change the basemap to create a black-and-white map
```{r geo_2, message=FALSE, warning=FALSE, paged.print=FALSE}
leaflet() %>% addTiles() %>% addMarkers(lat = as.numeric(geo$lat),
                                        lng = as.numeric(geo$lon)) %>% 
  addProviderTiles(providers$Stamen.Toner)
```