
---
title: "ggplot solutions"


output:
  rmdformats::material:
    highlight: kate
    self_contained: no
    gallery : T
    lightbox: T
    
---

##### 
```{r knitr_init, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, paged.print=FALSE}
rm(list = ls())
#1-7
echo_p1 = T
#8-17
echo_p2 = T
#18-22
echo_p3 = T
#23-33
echo_p4 = T
#34-40
echo_p5 = T
#41-48
echo_p6 = T
#49-51
echo_p7 = T
#52-54
echo_p8 = T
#55-57
echo_p9 = T
#58-61
echo_p10 = F
#62-64
echo_p11 = F
#65-70
echo_p12 = T

library(knitr)
library(rmdformats)


## Global options
opts_chunk$set(echo=F,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=100)

packages <- c('data.table','ggplot2','purrr','scales','dplyr', 'forcats','ggridges','knitr','kableExtra','magrittr')
invisible(sapply(packages, function(x) library(x,character.only = T) ))

```


# DATASETS

## MTCARS \

***

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).



* A data frame with 32 observations on 11 variables. 
      + mpg	Miles/(US) gallon
      + cyl	Number of cylinders
      + mpg	Miles/(US) gallon
      + disp	Displacement (cu.in.)
      + hp	Gross horsepower
      + drat	Rear axle ratio
      + wt	Weight (1000 lbs)
      + qsec	1/4 mile time
      + vs	V/S
      + am	Transmission (0 = automatic, 1 = manual)
      + gear	Number of forward gears
      + carb	Number of carburetors

          
```{r}
?datasets::mtcars
```
\

** Tidy mtcars ** 
```{r echo=T }
# mtcars datasets is almost tidy
# we have just to coerce it into a data.table, setting keep.rownames=T to get
# the information stored in the rownames (the car name). But we we do that the
# default name is 'rn', so we have to change the'rn' to 'name'
data.table(datasets::mtcars,keep.rownames = T) %>% setnames('rn','name') -> mtcars
```


## DIAMONDS \

***

A dataset containing the prices and other attributes of almost 54,000 diamonds. The variables are as follows:



* A data frame with 53940 rows and 10 variables: 
      + price	price in US dollars
      + carat	weight of the diamond (0.2 - 5.01)
      + cut	quality of the cut (Fair, Good, Very Good, Premium, Ideal)
      + color	diamond colour, from J (worst) to D (best)
      + clarity	a measurement of how clear the diamond is (I1 (worst), SI1, SI2, VS1, VS2, VVS1, VVS2, IF (best))
      + x	length in mm (0 - 10.74)
      + y	width in mm (0 - 58.9)
      + z	depth in mm (0 - 31.8)
      + depth	total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43 - 79)
      + table	width of top of diamond relative to widest point (43 - 95)

## TOOLTHGROWH \

***
**The Effect of Vitamin C on Tooth Growth in Guinea Pigs**
The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).


* A data frame with 60 rows and 3 variables: 
      + len	(numeric) Tooth lenght
      + supp (factor)	Supplement type (VC or OJ).
      + dose (numeric) Dose in milligrams/day



          
          

# AESTHETICS 



## AESTHETICS VS ATTRIBUTES
#### Attributes are used to change the way something look (shape,color,alpha,etc).
#### Aesthetics are used to map variables into attribues.



#### 1. Plot this graph
###### Use mtcars dataset. Map wt to x and mpg to y (set the aesthetics in the ggplot function, not in the geometric function )
```{r echo=echo_p1}
ggplot(mtcars,aes(x= wt, y = mpg)) + 
  geom_point()
```

#### 2. Increase the size of points
###### (copy and paste the ggplot code and set the size to 5)

```{r echo=echo_p1}
ggplot(mtcars,aes(x= wt, y = mpg, size = 5)) + 
  geom_point()
```

#### 3. Change the color of points
###### (copy and paste the ggplot code and set the color to  'hotpink')
###### there's an immense amount of colors for you to choose , click to see all of them [r colors](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)

```{r echo=echo_p1}
ggplot(mtcars,aes(x= wt, y = mpg, size = 5, color = 'hotpink')) + 
  geom_point()
```

#### 4. Up to know we have changed attributes of the plot. 
##### Now we are going to add a new dimension to the plot. Right now the plot has two dimension (mpg and wt). 
##### To add a new dimension, or better said, a new aesthetic, map the variable gear to the attribute color

```{r echo=echo_p1}
ggplot(mtcars,aes(x= wt, y = mpg, size = 5, color = gear)) + 
  geom_point()
```

## CATEGORICAL AESTHETICS

#### 5. CATEGORICAL AESTHETICS. Force the new aesthetic to be a categorical variable, a factor. 
##### To do so, use the function 'factor' when setting the color aesthetic
```{r echo=echo_p1}
ggplot(mtcars,aes(x= wt, y = mpg, size = 5, color = factor(gear))) + 
  geom_point()
```

## SHARED AESTHETICS VS GEOM-SPECIFIC AESTHETICS
#### 6. Add a new geometric to plot a loess model sharing all the aesthetics set 
##### use the geom_smooth function
```{r echo=echo_p1}
ggplot(mtcars,aes(x= wt, y = mpg, size = 5, color = factor(gear))) + 
  geom_point() +
  geom_smooth()
```

#### 7. Geom-specific aesthetic
##### As we can see in the plot above We have made a model for each year, and the size of the line is so big. 
##### This is because all the aesthetics set are shared aesthetics. 
##### So now, keep x and y as shared aesthetics but make size and color geom-specific aesthetics which only applies to the scatter plot.
```{r echo=echo_p1}
ggplot(mtcars,aes(x= wt, y = mpg )) + 
  geom_point(aes(size = 5, color = factor(gear))) +
  geom_smooth()
```


#### 8. Plot his graph
##### Use mtcars dataset. Map cyl to x and hp to y (set the aesthetics in the ggplot function, not in the geometric function )

```{r echo=echo_p2}
mtcars %>% ggplot(aes(cyl,hp)) + geom_point()
```

#### 9. Aesthetics and attributes
##### Add a geom-specific aesthetic to map the variable gear with the attribute color. 
##### And add also the attribute size. Set size = 3 

```{r echo=echo_p2}
mtcars %>% ggplot(aes(cyl,hp)) + geom_point(aes(color = factor(gear)), size = 3)
```

#### 10.Jitter position
##### Set the position argument to 'jitter'
```{r echo=echo_p2}
mtcars %>% ggplot(aes(cyl,hp)) + geom_point(aes(color = factor(gear)), size = 3, position = 'jitter')
```

#### 11. Change the default arguments of jitter position
##### To do so you need to use the jitter position (or the jitter geom)
##### set height = 0 and width =0.2 

```{r echo=echo_p2}
mtcars %>% ggplot(aes(cyl,hp)) + 
  geom_point(aes(color = factor(gear)), size = 3, position = position_jitter(width = 0.2,height = 0))
```
#### 12. Plot this graph
##### Use mtcars dataset. Map gear to x and use bar geometric
```{r echo=echo_p2}
 mtcars %>% ggplot(aes(x=gear)) + geom_bar()
```

#### 13. Add a new aesthetic. 
##### Map carb to fill aesthetic

```{r echo=echo_p2}
 mtcars %>% ggplot(aes(x=gear,fill=carb)) + geom_bar()
```

#### 14. Factor aesthetic 
##### We need to force carb to be a factor for the aesthetic to be visible

```{r echo=echo_p2}
 mtcars %>% ggplot(aes(x=gear,fill=factor(carb))) + geom_bar()
```

#### 15. Check default position
##### Set position argument to 'stack' to check that nothing changes

```{r echo=echo_p2}
 mtcars %>% ggplot(aes(x=gear,fill=factor(carb))) + geom_bar(position='stack')
```

#### 16. Fill position
##### Create a 100% stacked barplot by changing the position to 'fill'

```{r echo=echo_p2}
 mtcars %>% ggplot(aes(x=gear,fill=factor(carb))) + geom_bar(position='fill')
```

#### 17. Dodge position
##### Preserves the vertical position of an geom while adjusting the horizontal position.
```{r echo=echo_p2} 
 mtcars %>% ggplot(aes(x=gear,fill=factor(carb))) + geom_bar(position='dodge') 
```

#### 18. Plot this graph
##### Use mtcars dataset. Map disp to x mpg to y. You have to use two geoms
```{r echo=echo_p3} 
 ggplot(data=mtcars,aes(x=disp,y=mpg)) + geom_point() + geom_smooth()
```


#### 19. Attributes and Aesthetics
##### Aesthethics. Use the color aesthetic to add the variable gear to the plot
##### Attributes Change the color of the line to black and increase the size of points to 3
##### Remove standard error (shaded area) (hint geom_smooth has a parameter called se) 
```{r echo=echo_p3} 
 ggplot(data=mtcars,aes(x=disp,y=mpg)) + geom_point(aes(color=factor(gear)),size=5) + geom_smooth(se=F,color='black')
```

#### 20. Scales
##### Change the color scale manually. Coerce the colors to be ???red???, ???green??? and ???blue???to 3


```{r echo=echo_p3} 
 ggplot(data=mtcars,aes(x=disp,y=mpg)) + geom_point(aes(color=factor(gear),size=5)) + geom_smooth(se=F,color='black') +
  scale_color_manual(values = c('red','green','blue'))
```

#### 21. Add a new aesthetic
##### Use the size aesthetic to add the categorical variable carb
```{r echo=echo_p3} 
 ggplot(data=mtcars,aes(x=disp,y=mpg)) + geom_point(aes(color=factor(gear),size=factor(carb))) + geom_smooth(se=F,color='black') +
  scale_color_manual(values = c('red','green','blue'))
```

#### 22. Highlight the difference of carb
##### Change the range of the size scale, set it to 1-10

```{r echo=echo_p3} 
 ggplot(data=mtcars,aes(x=disp,y=mpg)) + geom_point(aes(color=factor(gear),size=factor(carb))) + geom_smooth(se=F,color='black') +
  scale_color_manual(values = c('red','green','blue'))+
  scale_size_discrete(range = c(1,10))
```




# GEOMETRICS 

#### 23. Plot this graph
##### Use diamonds dataset. Map price to x and create something like and histogram but plotting an area.

```{r echo=echo_p4} 
 ggplot(diamonds,aes(x=price)) + geom_area(stat='bin')
```

#### 24. Plot a density of the price 
```{r echo=echo_p4} 
 ggplot(diamonds,aes(x=price)) + geom_density()
```
#### 25. Plot the count instead of the density

```{r echo=echo_p4} 
 ggplot(diamonds,aes(x=price)) + geom_density(aes(y=..count..))
```

#### 26. Plot the count but using the triangular kernel
```{r echo=echo_p4} 
 ggplot(diamonds,aes(x=price)) + geom_density(kernel = 'rectangular',aes(y=..count..))
```

#### 27.Use a dotplot
```{r echo=echo_p4} 
 ggplot(diamonds,aes(x=price)) + geom_dotplot()
```

#### 28.We must reduce the dotsize. Set the attribute dotsize = 100/nrow(diamonds)
###### When the dataset is so big, the dotplot do not add information
```{r echo=echo_p4} 
 ggplot(diamonds,aes(x=price)) + geom_dotplot(dotsize = 100/nrow(diamonds))
```

#### 29.In this case we prefer to use a simple histogram
```{r echo=echo_p4} 
 ggplot(diamonds,aes(x=price)) + geom_histogram()
```

#### 30.Or a freqpoly (an histogram plotted using a line)
```{r echo=echo_p4} 
 ggplot(diamonds,aes(x=price)) + geom_freqpoly()
```

#### 31. Plot the count of each color using diamonds dataset
```{r echo=echo_p4} 
ggplot(diamonds,aes(cut)) + geom_bar()
```
#### 32. Add the aesthetic fill to map the variable color
```{r echo=echo_p4} 
ggplot(diamonds,aes(cut)) + geom_bar(aes(fill=factor(color)))
```

#### 33. Create a 100% stacked bar changing the default position
```{r echo=echo_p4} 
ggplot(diamonds,aes(cut)) + geom_bar(aes(fill=factor(color)), position='fill')
```

#### 34. Plot this famouse graph. 
##### Use the diamonds dataset and make a scatter plot of price vs carat
```{r echo=echo_p5}
ggplot(diamonds,aes(carat,price)) + geom_point() 
```

#### 35. Limit the x axis to 3 using a scale function
```{r echo=echo_p5}
ggplot(diamonds,aes(carat,price)) + geom_point() + scale_x_continuous(limits = c(NA,3))
```
#### 36. Use jitter instead of points and set height = 0.5,width = 0.5. 
###### You must also set the attribute alpha to 0.1 to avoid overlapping
```{r echo=echo_p5}
ggplot(diamonds,aes(carat,price)) + geom_jitter(height = 0.5,width = 0.5,alpha=0.1) + scale_x_continuous(limits = c(NA,3))
```

#### 37. Add quantiles 25%,50% and 75%

```{r echo=echo_p5}
ggplot(diamonds,aes(carat,price)) + geom_jitter(height = 0.5,width = 0.5,alpha=0.1) + geom_quantile(size = 2)+ scale_x_continuous(limits = c(NA,3))
```

#### 38. Limit the y axis by using the maximum price in the diamonds dataset.
```{r echo=echo_p5}
ggplot(diamonds,aes(carat,price)) + geom_jitter(height = 0.5,width = 0.5,alpha=0.1) + geom_quantile(size = 2)+
  scale_y_continuous(limits = c(NA,max(diamonds$price)))+ 
  scale_x_continuous(limits = c(NA,3))
```

#### 39.Visualize the distribution of price and carat using a rug plot (set alpha = 0.01 to avoid overlapping) 
```{r echo=echo_p5}
ggplot(diamonds,aes(carat,price)) + geom_jitter(height = 0.5,width = 0.5,alpha=0.1) + geom_quantile(size = 2)+
geom_rug(alpha=0.01) + 
  scale_y_continuous(limits = c(NA,max(diamonds$price)))+
  scale_x_continuous(limits = c(NA,3))
```

#### 40. Add a smooth model to see the relation between price and carat
##### Set attribute color = 'red' and zie = 2

```{r echo=echo_p5}
ggplot(diamonds,aes(carat,price)) + geom_jitter(height = 0.5,width = 0.5,alpha=0.1) + geom_quantile(size = 2)+
geom_rug(alpha=0.01) + geom_smooth(size=2,color='red')+
  scale_y_continuous(limits = c(NA,max(diamonds$price)))+
  scale_x_continuous(limits = c(NA,3))
```



#### 41. Plot thi graph (based on the previously created data.frame)
```{r echo=echo_p6}
diamonds <- as.data.table(diamonds)
df = diamonds[,.(price=mean(price)),by=color]
ggplot(df,aes(color,price)) + geom_bar(stat='identity')
```

#### 42. Add the color aesthetic
```{r echo=echo_p6}
df = diamonds[,.(price=mean(price)),by=color]
ggplot(df,aes(color,price)) + geom_bar(stat='identity',aes(fill=factor(color)))
```


#### 43. We can get more information without creating an auxiliar dataset.
###### Plot this graph using the diamonds dataset and a boxplot geometric
```{r echo=echo_p6}
ggplot(diamonds,aes(color,price)) + geom_boxplot(aes(fill=factor(color)))
```

#### 44. Based on the following dataset. Create a dotplot
###### Remember that the argument binaxis must be set to 'y'
```{r echo = echo_p6}
diamonds = as.data.table(diamonds)
df = diamonds[,.(price=mean(price)),by=.(cut,color,clarity)]
ggplot(df,aes(color,price))+ geom_dotplot(binaxis = 'y', dotsize = 0.7, aes(fill=color), stackdir = 'center')
```

#### 45. To create a plot with the same information, we don't need to create an auxliar dataset
###### Plot this graph using the diamonds dataset and a violin geometric
```{r echo=echo_p6}
ggplot(diamonds,aes(color,price)) + geom_violin(aes(fill=factor(color)))
```





#### 46. Plot this graph. As you can see, it has been made using two categorical variables
```{r echo=echo_p6}
diamonds %>% ggplot() + geom_jitter(aes(clarity,cut))
```

#### 47. Add two new aesthetics
###### Map the carat variable in the alpha attribute and the variable price in the color attribute
```{r echo=echo_p6}
diamonds %>% ggplot() + geom_jitter(aes(clarity,cut,alpha=carat,color=price)) 
```

#### 48. Change the range of the alpha aesthetic using a scale function 
###### Change the default theme, use theme_minimal
```{r echo=echo_p6}
diamonds %>% ggplot() + geom_jitter(aes(clarity,cut,alpha=carat,color=price)) +
  scale_alpha_continuous(range = c(0,0.7))+
  theme_minimal()
```

#### 49. Plot this graph. 
###### Use diamonds dataset, map carat to x and price to y and use hex geometric (set alpha to 0.8)
```{r echo=echo_p7}
diamonds %>% ggplot(aes(carat,price)) + geom_hex(alpha=0.8)
```


#### 50. Add a contours. 
###### set color = 'red' to make it more visible
```{r echo=echo_p7}
diamonds %>% ggplot(aes(carat,price)) + geom_hex(alpha=0.8) + geom_density2d(color='red')
```


#### 51. Change geom_text with geom_point 
###### set alpha  to 0.005 to be ableo to see somthing
```{r echo=echo_p7}
diamonds %>% ggplot(aes(carat,price)) + geom_point(alpha=0.005) + geom_density2d(color='red')
```

#### 52. Using the following dataset. Plot a line to know the trend of the unenmployment 
```{r comment='##', echo=echo_p8}
# CREATE A US ECONOMIC DATASET
df <- as.data.table(datasets::longley)
df[,year := as.integer(rownames(datasets::longley))]
colnames(df) <- gsub('\\.','_',tolower(colnames(df)))
# plot the line
ggplot(df)+ geom_line(aes(y=unemployed,x=year))
```

#### 53. Plot an area instead
```{r comment='##', echo=echo_p8}
ggplot(df)+ geom_area(aes(y=unemployed,x=year))
```

#### 54. Plot an step line
```{r comment='##', echo=echo_p8}
ggplot(df)+ geom_step(aes(y=unemployed,x=year))
```


#### 55. Create the model an plot a crossbar
```{r comment='##', echo=echo_p9}
diamonds <- as.data.table(ggplot2::diamonds)
diamonds %>% 
  .[,unique(color)] %>% 
  as.character() %>% 
  map(~diamonds[color==.]) %>% 
  map(~lm(formula = log(price)~log(carat),data=.)) %>% 
  map(coefficients) %>% 
  map(as.data.frame) %>% 
  rbindlist() %>% 
  as.data.table() %>% 
  setnames(.,names(.),'coef') %>% 
  .[,type := rep(c('intercept','log_carat'),diamonds[,uniqueN(color)])] %>% 
  .[,color:=unlist(map(as.character(diamonds[,unique(color)]),~rep(.,2)))] %>% 
  dcast(color~type,value.var='coef') %>% 
  merge(diamonds,.,by='color') %>% 
  .[,price_hat := exp(log_carat * log(carat) + intercept)]->diamonds
#ggplot(diamonds,aes(price,price_hat))+ geom_point(alpha=0.3,aes(color=color))+geom_smooth()

df <- diamonds[,.(fit=mean(price_hat),sd = mean(sqrt((price_hat-price)**2))), by= color]

ggplot(df,aes(color,fit,ymin=fit-sd,ymax=fit+sd))+
  geom_crossbar(aes(fill=color))

```

#### 56. Plot this graph
##### Use geom bar (stat ='identity') and add an error bar
```{r echo=echo_p9}
ggplot(df,aes(color,fit,ymin=fit-sd,ymax=fit+sd))+
    geom_bar(stat='identity',aes(fill=color)) +
    geom_errorbar()
```



#### 57. Use a point-range instead of an error bar
```{r echo=echo_p9}
ggplot(df,aes(color,fit,ymin=fit-sd,ymax=fit+sd))+
    geom_bar(stat='identity',aes(fill=color)) +
    geom_pointrange()
```

# FACETS
#### 58. Plot this graph
```{r echo = echo_p10}
ggplot(diamonds,aes(price)) + geom_histogram()
```

#### 59. Split the plot using the categorical variable color
```{r echo = echo_p10}
ggplot(diamonds,aes(price)) + geom_histogram() + facet_wrap(~color)
```


#### 60. Create a grid of plot using the variables color in columns and cut in rows.
```{r echo = echo_p10}
ggplot(diamonds,aes(price)) + geom_histogram() +  facet_grid(cut~color)
```

#### 61. Create a grid of plot using the variables color in columns and cut in rows.
```{r echo = echo_p10}
ggplot(diamonds,aes(price)) + geom_histogram() +  facet_grid(cut~color,scales='free')
```

# COORDINATES
#### 62. Plot this graph. 
###### Hints theta = 'y', x=''
```{r echo = echo_p11}
ggplot(diamonds, aes(x = '', fill = factor(cut))) +
  geom_bar(width = 1,position='fill')+ coord_polar(theta = "y")
```



#### 63. Stat transformation using coordinate
###### sqrt in x and log10 in price
```{r echo = echo_p11}
ggplot(diamonds,aes(carat,price)) + geom_point(alpha=0.1) + geom_smooth() +
  coord_trans(x='sqrt',y='log10') + labs(x='sqrt(carat)',y='log10(price)')
```

#### 64. Flipped violins
```{r echo = echo_p11}
ggplot(diamonds,aes(color,price,fill=factor(color))) + geom_violin(alpha=0.3) +
  coord_flip()
```

# THEMES
#### 65. Based on this plot. Add a theme to make it look more classic.
```{r echo = echo_p12}
ggplot(ToothGrowth,aes(factor(dose),len,fill=factor(dose))) +  
  geom_violin(trim=FALSE)+
  stat_summary(fun.data=function(x)c(y=mean(x),ymin=mean(x)-sd(x),ymax=mean(x)+sd(x))) +
  scale_fill_brewer(palette="Blues") ->gg
gg + theme_classic()
```

#### 66. Add the title "Tooth Length by Dose" and Subtitle "Vitamin C Complex"
```{r echo = echo_p12}
gg + theme_classic() + ggtitle("Tooth Length by Dose", subtitle = 'Vitamin C Complex')
```
  
#### 67. Add a new Label for X Axis "Length", a new one for Y Axis "Dose (mg)", and a new title for the legend "Dose"
```{r echo = echo_p12}
gg + theme_classic() + ggtitle("Tooth Length by Dose", subtitle = 'Vitamin C Complex') +
  labs(y = 'Length', x = 'Dose (mg)', fill = 'Dose')
```


#### 68. Plot the title and subtitle in the center
```{r echo = echo_p12}
gg + theme_classic() + ggtitle("Tooth Length by Dose", subtitle = 'Vitamin C Complex') +
  labs(y = 'Length', x = 'Dose (mg)', fill = 'Dose') +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))
```


#### 69. Increase the base size and set the base family = 'serif'
```{r echo = echo_p12}
gg + theme_classic(base_size = 15, base_family = 'serif') + ggtitle("Tooth Lenght by Dose", subtitle = 'Vitamin C Complex') +
  labs(y = 'Length', x = 'Dose (mg)', fill = 'Dose') +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))
```


#### 70. Change the aspect ratio of the coordinate system. Fix the ratio to 1/15
```{r echo = echo_p12}
gg + theme_classic(base_size = 15, base_family = 'serif') + ggtitle("Tooth Lenght by Dose", subtitle = 'Vitamin C Complex') +
  labs(y = 'Length', x = 'Dose (mg)', fill = 'Dose') +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))+
  coord_fixed(ratio = 1/15)
```




